Ornaments as indicators of social changes in northeastern Taiwan before and after the European colonial period

### 2020-04-24 analysis/paper/paper.Rmd
add suggested references from R2

Examination of the archaeological record at the peripheries of colonial activities can show how remote Indigenous groups were affected by major European colonial processes [@Trabert2017]. Kiwulan in northeastern Taiwan is an exceptional case study as an East Asian location that was relatively isolated and peripheral, and yet connected by regional and global trade networks. Kiwulan provides valuable insights into the discussion of indirect colonial influence on local societies living beyond the borders of direct European colonial occupation. The frequency and spatial distribution of personal ornaments at Kiwulan present three distinct patterns during different dominant culture interaction periods. The greater amount and diversity of ornament types during the European period reflects an increasing use in ornaments in a colonial context. Before European contact, ornaments were traded into local Indigenous societies via the regional exchange network with Chinese merchants and viewed as prestige goods in local Indigenous culture according to their distribution in the archaeological contexts. After the arrival of the Europeans, the exotic and powerful image carried by those ornaments may have intensified, further signaling wealth and privileged trading connections among the inhabitants of Kiwulan. This may have stimulated more competition between aggrandizing individuals for prestige and wealth accumulation at Kiwulan, which might have resulted in an increase in social inequality. This might also indicate an act of intentional resistance to the intrusion of the Europeans by using more ornaments that are symbolic of cultural tradition.

>>>>

Examination of the archaeological record at the peripheries of colonial activities can show how remote Indigenous groups were affected by major European colonial processes [@Trabert2017], and similar impacts can be also seen in modern societies today when adopting western products [@AlGhanim2017; @Jane2015]. Kiwulan in northeastern Taiwan is an exceptional case study as an East Asian location that was relatively isolated and peripheral, and yet connected by regional and global trade networks. Kiwulan provides valuable insights into the discussion of indirect colonial influence on local societies living beyond the borders of direct European colonial occupation. The frequency and spatial distribution of personal ornaments at Kiwulan present three distinct patterns during different dominant culture interaction periods. The greater amount and diversity of ornament types during the European period reflects an increasing use in ornaments in a colonial context. Before European contact, ornaments were traded into local Indigenous societies via the regional exchange network with Chinese merchants and viewed as prestige goods in local Indigenous culture according to their distribution in the archaeological contexts. After the arrival of the Europeans, the exotic and powerful image carried by those ornaments may have intensified, further signaling wealth and privileged trading connections among the inhabitants of Kiwulan. This may have stimulated more competition between aggrandizing individuals for prestige and wealth accumulation at Kiwulan, which might have resulted in an increase in social inequality. This might also indicate an act of intentional resistance to the intrusion of the Europeans by using more ornaments that are symbolic of cultural tradition.
